A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-12 docs/icccn2009/paper.tex
ICCCN 2009 Paper Update;
Integrating Kleber's review. All changes to this
revision were submitted by Kleber via email.

git-svn-id: svn://localhost/netbeams/trunk@367 559edde2-a4d4-44e0-96fc-4035e40d7396

Should describe DSP architecture. Only mention OSGi in last subsection
and explain that components can be mapped to bundles. Currently no
distinction between Platform and Framework. Necessary?

@@

Figure \ref{FIG_DSP} depicts the basic architecture of the
DSP. We refer to the plug-and-play modules as DSP Components. DSP
Components are self-contained modules that can be added and removed to
a DSP installation at runtime. The basic communication paradigm within
a DSP is a message. DSP Components can exchange messages of different
types. We distinguish between Data Producers (DP) and Data Consumers
(DP) within the scope of a DSP. DP and DC can be seen as roles that a
DSP Component may have. It is possible for a DSP Component to act in
the role of DC and DP at the same time.

@@

By implementing the above Java interface, the resuling code
effectively becomes a DSP Component that can be deployed at runtime. A
component is identified by its type. Only one component of a certain
type per DSP is permissible. A string serves as a type identifier that
much be unique across all DSP Components. Upon deploying a DSP
Component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 5) and stopped (line
6) multiple time (TODO: SHOULDN'T THERE BE ALSO A WAY TO UNDEPLOY A
COMPONENT?) Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (ine
7). The message structure will be explained in the following section.

>>>>

Sensor networks, especially of environmental sensors, are normally deployed
over distributed locations where communication infrastructure are not always 
reliable or, at least, not countinuously available. Additionally, different
sensors and communication protocols might be available in throughout the network.

As the network grows in size and complexity scalability, reliabilty and maintainability
become critical issues. Additional components 
might be added to the network, the network configuration might be altered or 
new versions of components might be deployed. The sensor network must be able to 
support such adjustments without significant overhead for network administrators and
the components in the network.

From system development and integration point of view, interfaces among components and
between components and the network should be simple. Complex interaction can be 
constructed on top of these simple interfaces ( a good example of this strategy is the 
world-wide-web).

In the following we discuss how our proposed architechure addresses those issues.

@@

Figure \ref{FIG_DSP} depicts the basic architecture of the
DSP. We refer to the plug-and-play modules as DSP Components. DSP
Components are self-contained modules that can be added and removed to
a DSP installation at runtime. The basic communication paradigm within
a DSP is a message. DSP Components can exchange messages of different
types. Actually, all interactions between DSP Components are implemented
through messages yielding simple but powerful communication instructure.
We distinguish between Data Producers (DP) and Data Consumers
(DC) within the scope of a DSP. DP and DC can be seen as roles that a
DSP Component may have. It is possible for a DSP Component to act in
the role of DC and DP at the same time. Messages in DSPs are routed based 
on rules. Rules can be statically or dynamically configured. Rules can use 
different source information when routing a message: producers or consumers of the message, 
the message itself or specific configuration provided by the administrators.

@@

By implementing the above Java interface, the resuling code
effectively becomes a DSP Component that can be deployed at runtime. A
component is identified by its type. Only one component of a certain type per DSP 
is permissible. A type identifier much be unique across all DSP Components in DSP. 
Upon deploying a DSP Component, an initialization function is called (line 4). Once a
component is initialized, it may be started (line 5) and stopped (line
6) multiple time. Whenever the DSP wishes to pass a message (that was
generated by some DP), it invokes the \texttt{deliver()} method (ine
7). The message structure will be explained in the following section.
One important feature of the DSP is the ability to redeploy (maybe with a new version) at 
runtime. This allows DSPs to be updated remotely. This process is managed through
messages sent to a special DSP Component which represents the DSP itself.
